---
title: "kappa"
author: "Kev Sheldrake"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = FALSE, message = FALSE)
```

```{r}
if ("xfun" %in% row.names(installed.packages()) == FALSE) {
  install.packages("xfun")
}

xfun::pkg_attach(
  c(
    "tidyverse",
    "magrittr",
    "bayesplay",
    "knitr",
    "effsize",
    "weights",
    "kableExtra",
    "LaplacesDemon"
  ),
  install = TRUE
)
```
# Load the data
```{r load the data}
yesesvi.df <- readr::read_csv("yesesvi-diachronic-analysis.csv", skip = 1, col_names = TRUE, show_col_type = FALSE)
kev.df <- readr::read_csv("kev-diachronic-analysis.csv", col_names = TRUE, show_col_type = FALSE)
yesesvi.syn.df <- readr::read_csv("yesesvi-synchronic-analysis.csv", skip = 1, col_names = TRUE, show_col_type = FALSE)
kev.syn.df <- readr::read_csv("kev-synchronic-analysis.csv", col_names = TRUE, show_col_type = FALSE)
```
# Recode the data
```{r recode data}
yesesvi.df <- yesesvi.df %>% dplyr::mutate(`#` = recode(`#`, '22' = '22.1', '22a' = '22.2', '22b' = '22.3', '15' = '15.2', '15a' = '15.1'))
yesesvi.syn.df <- yesesvi.syn.df %>% dplyr::mutate(`#` = recode(`#`, '22' = '22.1', '22a' = '22.2', '22b' = '22.3', '15' = '15.2', '15a' = '15.1'))
yesesvi.df <- yesesvi.df %>% dplyr::mutate(Moment = as.character(Moment))
yesesvi.df <- yesesvi.df %>% dplyr::mutate(Moment = recode(Moment, '6' = '7'))
kev.df <- kev.df %>% dplyr::mutate(`#` = as.character(`#`))
kev.syn.df <- kev.syn.df %>% dplyr::mutate(`#` = as.character(`#`))
kev.syn.df <- kev.syn.df %>% dplyr::filter(ISUnum > 0)
kev.syn.df <- kev.syn.df %>% dplyr::mutate(ISUnum = as.character(ISUnum))
```
# Remove experimenter utterances
```{r remove experimenter utterances}
yesesvi.df <- yesesvi.df %>% dplyr::filter(! `#` %in% c('10', '21', '23', '29', '12', '14', '16', '19', '39', '27', '31', '33', '35', '37', '25'))
kev.df <- kev.df %>% dplyr::filter(! `#` %in% c('10', '21', '23', '29', '12', '14', '16', '19', '39', '27', '31', '33', '35', '37', '25'))
yesesvi.syn.df <- yesesvi.syn.df %>% dplyr::filter(! `#` %in% c('10', '21', '23', '29', '12', '14', '16', '19', '39', '27', '31', '33', '35', '37', '25'))
kev.syn.df <- kev.syn.df %>% dplyr::filter(! `#` %in% c('10', '21', '23', '29', '12', '14', '16', '19', '39', '27', '31', '33', '35', '37', '25'))
```
# Calculate utterance totals
```{r get totals}
uttnums <- str_unique(c(yesesvi.df$`#`, kev.df$`#`))
uttnums.syn <- str_unique(c(yesesvi.syn.df$`#`, kev.syn.df$`#`))
```
# Calculate
```{r}
agree <- 0
expected <- 0
for (category in c(1,2,3,4,5,6,7)) {
  agree <- agree + length(intersect((kev.df %>% dplyr::filter(Moment == as.character(category)))$`#`, (yesesvi.df %>% dplyr::filter(Moment == as.character(category)))$`#`))
  expected <- expected + (length((kev.df %>% dplyr::filter(Moment == as.character(category)))$`#`) * length((yesesvi.df %>% dplyr::filter(Moment == as.character(category)))$`#`))
}
```

```{r}
agree.syn <- 0
expected.syn <- 0
for (category in c(1,2,3,4,5,6,7,8,9)) {
  agree.syn <- agree.syn + length(intersect((kev.syn.df %>% dplyr::filter(ISUnum == as.character(category)))$`#`, (yesesvi.syn.df %>% dplyr::filter(ISUnum == as.character(category)))$`#`))
  expected.syn <- expected.syn + (length((kev.syn.df %>% dplyr::filter(ISUnum == as.character(category)))$`#`) * length((yesesvi.syn.df %>% dplyr::filter(ISUnum == as.character(category)))$`#`))
}
```

```{r}
expfreq <- expected / length(uttnums)
kappa <- (agree - expfreq) / (length(uttnums) - expfreq)
expfreq.syn <- expected.syn / length(uttnums.syn)
kappa.syn <- (agree.syn - expfreq.syn) / (length(uttnums.syn) - expfreq.syn)
```
# Conclusions
For the diachronic analysis, Cohen's kappa = `r round(kappa, 2)`, which implies strong agreement and is typically considered adequate. For the synchronic analysis, Cohen's kappa = `r round(kappa.syn, 2)`, which implies moderate agreement and is typically considered adequate (McHugh, 2012).

The calculation is based on the description here: https://psych.unl.edu/psycrs/handcomp/hckappa.PDF

# References
McHugh, M. L. (2012). Interrater reliability: the kappa statistic. Biochemia medica, 22(3), 276-282.
